---
title: "K562 MYC Chromatin Screen"
output: html_document
---
#--------------------------------------------
# K562 MYC Chromatin Screen
#--------------------------------------------
K562 MYC-mNeonGreen cells were transduced with the human chromatin focussed sgRNA library. 
GFP low cells were enriched by FACS on days 3,4,5,6 post transduction and sequenced --> NEG fraction
Timepoint matched LIB control fractions were taken at each timepoint. 
MID fractions were also sorted and sequenced. 
Two replicates were performed for both clone 32 and clone 58. Therefore n = 4

BARtab v1.4 was used to extract and count sgRNAs from raw sequencing reads
Resulting counts tables are imported here for CRISPR screen analysis using edgeR.

---

# Setup

## import libraries
```{r, message=FALSE, warning=FALSE}
library(edgeR)
library(bartools)
library(tidyverse)
library(ggpointdensity)
library(viridis)
library(reshape2)
#library(MAGeCKFlute)
library(RColorBrewer)
```

## setup output dirs
```{r, message=FALSE, warning=FALSE}
results.dir <- file.path(getwd(), "edgeR/results")
dir.create(results.dir, showWarnings = F)

plots.dir <- file.path(getwd(), "edgeR/plots")
dir.create(plots.dir, showWarnings = F)
```

#------------------------------------------------
# Setup DGEList object
#------------------------------------------------

## import counts files
```{r}
raw.counts <- read.delim("./K562_temporal_chromatin_raw_counts.tsv", header = T, sep = "\t", row.names = 1)
colnames(raw.counts) <- gsub("X", "C", colnames(raw.counts))

# separate count table per clone
raw.counts.58 <- raw.counts[,grep("C58", colnames(raw.counts))]
colnames(raw.counts.58) <- c("C58.R2.D5.NEG","C58.R1.D6.NEG","C58.R1.D4.LIB","C58.R1.D4.NEG","C58.R2.D4.MID","C58.R2.D5.MID","C58.R2.D3.NEG","C58.R2.D6.NEG","C58.R1.D5.NEG","C58.R2.D4.NEG","C58.R2.D6.MID","C58.R2.D3.MID","C58.R1.D5.LIB","C58.R2.D3.LIB","C58.R1.D6.LIB","C58.R2.D6.LIB","C58.R2.D4.LIB","C58.R2.D5.LIB")

raw.counts.32 <- raw.counts[,grep("C32", colnames(raw.counts))]
colnames(raw.counts.32) <- c("C32.R2.D6.MID","C32.R1.D4.NEG","C32.R2.D4.MID","C32.R2.D4.NEG","C32.R2.D5.NEG","C32.R2.D5.MID","C32.R1.D5.NEG","C32.R2.D3.MID","C32.R2.D6.NEG","C32.R1.D6.NEG","C32.R2.D3.NEG","C32.R1.D4.LIB","C32.R2.D3.LIB","C32.R2.D5.LIB","C32.R1.D6.LIB","C32.R2.D6.LIB")

raw.counts <- cbind(raw.counts.32, raw.counts.58)
raw.counts <- na.omit(raw.counts)
```


## generate metadata
```{r}
metadata <- as.data.frame(str_split(colnames(raw.counts), pattern = "\\.", simplify = T), row.names = colnames(raw.counts))
colnames(metadata) <- c("Clone", "Replicate", "Day", "Group")

# fix gene names
genes <- as.data.frame(str_split(row.names(raw.counts), "_", simplify = T))
sgRNAs <- as.data.frame(str_split(genes[,3], "\\.", simplify = T))
genes$V4 <- sgRNAs$V2
colnames(genes) <- c("Symbol", "ENSEMBL", "ENTREZ", "Guide")
```

## create DGEList
```{r}
dge <- DGEList(counts = as.matrix(raw.counts), samples = metadata, group = metadata$Group, genes = genes, remove.zeros = T)
dge
```

## filter DGEList
```{r}
table(dge$samples$group)

# Filter sgRNAs and samples with low counts Need a CPM greater than 5 in 15 or more samples
# to keep sgRNAs
selr = rowSums(cpm(dge$counts) > 5) >= 15
# Need at least 100,000 reads to keep a given sample
selc = colSums(dge$counts) >= 1e+05
dge = dge[selr, selc]
dim(dge)
```

## Plot counts per sample & guide
```{r}
# Set up drug treatment colours
cols = as.numeric(dge$samples$group) + 1

barplot(colSums(dge$counts), las = 2, main = "Counts per index", col = cols, cex.names = 0.5,
cex.axis = 0.8)
barplot(rowSums(dge$counts), las = 2, main = "Counts per sgRNA", cex.names = 0.5, cex.axis = 0.8)

pdf(file.path(plots.dir, "K562_Chromatin_counts_per_sample_sgRNA.pdf"))
barplot(colSums(dge$counts), las = 2, main = "Counts per index", col = cols, cex.names = 0.5,
cex.axis = 0.8)
barplot(rowSums(dge$counts), las = 2, main = "Counts per sgRNA", cex.names = 0.5, cex.axis = 0.8)
dev.off()

```
#------------------------------------------------
# NEG sample analysis - Combined clones
#------------------------------------------------

## MDS plot
```{r}
dge = calcNormFactors(dge, method = "none")

# Make an MDS plot to visualise relationships between replicate samples Set up infection #
# colours
cols = as.numeric(factor(dge$samples$Group)) + 1
cols2 = as.numeric(dge$samples$Clone) + 1
cols3 = as.numeric(factor(dge$samples$Replicate)) + 1
plotMDS(dge, col = cols, main = "MDS Plot")
plotMDS(dge, col = cols, main = "MDS Plot", pch = 16)
plotMDS(dge, col = cols2, main = "MDS Plot")
plotMDS(dge, col = cols3, main = "MDS Plot")

pdf(file.path(plots.dir, "K562_Chromatin_MDS_plots.pdf"))
plotMDS(dge, col = cols, main = "MDS Plot")
plotMDS(dge, col = cols, main = "MDS Plot", pch = 16)
plotMDS(dge, col = cols2, main = "MDS Plot")
plotMDS(dge, col = cols3, main = "MDS Plot")
dev.off()
```

## Set up design matrix for GLM
We will use GLMs in edgeR in this case since there are more than 2 groups 
```{r}
group = as.factor(dge$samples$group)
replicate = as.factor(dge$samples$Replicate)
timepoint = as.factor(dge$samples$Day)
clone = as.factor(dge$samples$Clone)
contrasts = as.factor(paste(timepoint,group, sep = '.'))
des = model.matrix(~0+contrasts)
colnames(des) <- levels(contrasts)
des[1:5, ]
```

## Estimate dispersions
```{r}
xglm = estimateDisp(dge, des)
sqrt(xglm$common.disp)
```

## Plot BCVs versus abundance
```{r}
plotBCV(xglm, main = "BCV Plot")

pdf(file.path(plots.dir, "K562_Chromatin_BCV_plot.pdf"))
plotBCV(xglm, main = "BCV Plot")
dev.off()
```

## Fit negative bionomial GLM
```{r}
fit = glmQLFit(xglm, des)

# define contrasts of interest
my.contrasts <- makeContrasts(
  D3 = D3.NEG-D3.LIB,
  D4 = D4.NEG-D4.LIB,
  D5 = D5.NEG-D5.LIB,
  D6 = D6.NEG-D6.LIB,
  levels = des)

qlf.D3 <- glmQLFTest(fit, contrast=my.contrasts[,"D3"])
write.csv(topTags(qlf.D3, n = Inf), file = file.path(results.dir, "K562_Chromatin_32_58_NEG_D3_toptags.csv"))

qlf.D4 <- glmQLFTest(fit, contrast=my.contrasts[,"D4"])
write.csv(topTags(qlf.D4, n = Inf), file = file.path(results.dir, "K562_Chromatin_32_58_NEG_D4_toptags.csv"))

qlf.D5 <- glmQLFTest(fit, contrast=my.contrasts[,"D5"])
write.csv(topTags(qlf.D5, n = Inf), file = file.path(results.dir, "K562_Chromatin_32_58_NEG_D5_toptags.csv"))

qlf.D6 <- glmQLFTest(fit, contrast=my.contrasts[,"D6"])
write.csv(topTags(qlf.D6, n = Inf), file = file.path(results.dir, "K562_Chromatin_32_58_NEG_D6_toptags.csv"))

# LRT tests
#lrt = glmLRT(fit, coef = 3)
# Show top ranked sgRNAs
#topTags(lrt)
```

## plot smear for D6
```{r}
# Select sgRNAs with FDR < 0.001 and logFC <= -1 to highlight on plot
thresh = 0.01
lfc = 1
top4 = topTags(qlf.D6, n = Inf)
top4ids = rownames(top4$table[top4$table$FDR < thresh & top4$table$logFC >= lfc,])
# Plot logFC versus logCPM
plotSmear(qlf.D6, de.tags = top4ids, pch = 20, cex = 0.6, main = "Large sgRNA-seq screen: logFC vs logCPM")
abline(h = c(-1, 0, 1), col = c("dodgerblue", "yellow", "dodgerblue"), lty = 2)

pdf(file.path(plots.dir, "32_58_NEG_D6_smear_plot.pdf"))
plotSmear(qlf.D6, de.tags = top4ids, pch = 20, cex = 0.6, main = "Large sgRNA-seq screen: logFC vs logCPM")
abline(h = c(-1, 0, 1), col = c("dodgerblue", "yellow", "dodgerblue"), lty = 2)
dev.off()

```

#------------------------------------------------
# Gene set testing
#------------------------------------------------

## setup gene list for camera
```{r}
genesymbols = dge$genes[,1]
genesymbollist = list()
unq = unique(genesymbols)
unq = unq[!is.na(unq)]
for (i in unq) {
  sel = genesymbols == i & !is.na(genesymbols)
  if (sum(sel) > 3){
    genesymbollist[[i]] = which(sel)
  }
}

write(unq, file = file.path(results.dir, "chrom_lib_genes.txt"))
```

## Run camera for all genes
```{r}
camera.res.D3 = camera(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D3"])
camera.res.D3[camera.res.D3$FDR < 0.05,]

camera.res.D4 = camera(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D4"])
camera.res.D4[camera.res.D4$FDR < 0.05,]

camera.res.D5 = camera(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D5"])
camera.res.D5[camera.res.D5$FDR < 0.05,]

camera.res.D6 = camera(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D6"])
camera.res.D6[camera.res.D6$FDR < 0.05,]
```

## Run roast for all genes
```{r}
roast.res.D3 = roast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D3"])
roast.res.D3[roast.res.D3$FDR < 0.1,] %>% filter(Direction == "Up")

roast.res.D4 = roast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D4"])
roast.res.D4[roast.res.D4$FDR < 0.1,] %>% filter(Direction == "Up")

roast.res.D5 = roast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D5"])
roast.res.D5[roast.res.D5$FDR < 0.1,] %>% filter(Direction == "Up")

roast.res.D6 = roast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D6"])
roast.res.D6[roast.res.D6$FDR < 0.1,] %>% filter(Direction == "Up")
```

## Run mroast for all genes
```{r}
mroast.res.D3 <- mroast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D3"])
mroast.res.D3[mroast.res.D3$FDR < 0.1,] %>% filter(Direction == "Up") %>% filter(PropUp > 0.4)

mroast.res.D4 <- mroast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D4"])
mroast.res.D4[mroast.res.D4$FDR < 0.1,] %>% filter(Direction == "Up") %>% filter(PropUp > 0.4)

mroast.res.D5 <- mroast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D5"])
mroast.res.D5[mroast.res.D5$FDR < 0.1,] %>% filter(Direction == "Up") %>% filter(PropUp > 0.4)

mroast.res.D6 <- mroast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D6"])
mroast.res.D6[mroast.res.D6$FDR < 0.1,] %>% filter(Direction == "Up") %>% filter(PropUp > 0.4)

```

## Run fry for all genes
```{r}
fry.res.D3 <- fry(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D3"], sort = "directional")
fry.res.D3[fry.res.D3$FDR < 0.3,] %>% filter(Direction == "Up")

fry.res.D4 <- fry(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D4"], sort = "directional")
fry.res.D4[fry.res.D4$FDR < 0.3,] %>% filter(Direction == "Up")

fry.res.D5 <- fry(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D5"], sort = "directional")
fry.res.D5[fry.res.D5$FDR < 0.3,] %>% filter(Direction == "Up")

fry.res.D6 <- fry(xglm, index = genesymbollist, des, contrast = my.contrasts[,"D6"], sort = "directional")
fry.res.D6[fry.res.D6$FDR < 0.3,] %>% filter(Direction == "Up")

```

## write gene set results to file
```{r}
# camera
write.table(camera.res.D3, file = file.path(results.dir, "32_58_D3_NEG_camera_res.csv"), sep = ",")
write.table(camera.res.D4, file = file.path(results.dir, "32_58_D4_NEG_camera_res.csv"), sep = ",")
write.table(camera.res.D5, file = file.path(results.dir, "32_58_D5_NEG_camera_res.csv"), sep = ",")
write.table(camera.res.D6, file = file.path(results.dir, "32_58_D6_NEG_camera_res.csv"), sep = ",")

# roast
write.table(roast.res.D3, file = file.path(results.dir, "32_58_D3_NEG_roast_res.csv"), sep = ",")
write.table(roast.res.D4, file = file.path(results.dir, "32_58_D4_NEG_roast_res.csv"), sep = ",")
write.table(roast.res.D5, file = file.path(results.dir, "32_58_D5_NEG_roast_res.csv"), sep = ",")
write.table(roast.res.D6, file = file.path(results.dir, "32_58_D6_NEG_roast_res.csv"), sep = ",")

# mroast
write.table(mroast.res.D3, file = file.path(results.dir, "32_58_D3_NEG_mroast_res.csv"), sep = ",")
write.table(mroast.res.D4, file = file.path(results.dir, "32_58_D4_NEG_mroast_res.csv"), sep = ",")
write.table(mroast.res.D5, file = file.path(results.dir, "32_58_D5_NEG_mroast_res.csv"), sep = ",")
write.table(mroast.res.D6, file = file.path(results.dir, "32_58_D6_NEG_mroast_res.csv"), sep = ",")

# fry
write.table(fry.res.D3, file = file.path(results.dir, "32_58_D3_NEG_fry_res.csv"), sep = ",")
write.table(fry.res.D4, file = file.path(results.dir, "32_58_D4_NEG_fry_res.csv"), sep = ",")
write.table(fry.res.D5, file = file.path(results.dir, "32_58_D5_NEG_fry_res.csv"), sep = ",")
write.table(fry.res.D6, file = file.path(results.dir, "32_58_D6_NEG_fry_res.csv"), sep = ",")

```

#------------------------------------------------
# PLOTS
#------------------------------------------------

## get sigificant genes from camera roast mroast and fry analysis
```{r}
hits <- unique(
  c(
  rownames(camera.res.D3[camera.res.D3$FDR < 0.01,]),
  rownames(camera.res.D4[camera.res.D4$FDR < 0.01,]),
  rownames(camera.res.D5[camera.res.D5$FDR < 0.01,]),
  rownames(camera.res.D6[camera.res.D6$FDR < 0.01,]),
  
  rownames(roast.res.D3[roast.res.D3$FDR < 0.01,]),
  rownames(roast.res.D4[roast.res.D4$FDR < 0.01,]),
  rownames(roast.res.D5[roast.res.D5$FDR < 0.01,]),
  rownames(roast.res.D6[roast.res.D6$FDR < 0.01,]),
  
  rownames(mroast.res.D3[mroast.res.D3$FDR < 0.01,]),
  rownames(mroast.res.D4[mroast.res.D4$FDR < 0.01,]),
  rownames(mroast.res.D5[mroast.res.D5$FDR < 0.01,]),
  rownames(mroast.res.D6[mroast.res.D6$FDR < 0.01,]),
  
  rownames(fry.res.D3[fry.res.D3$FDR < 0.01,]),
  rownames(fry.res.D4[fry.res.D4$FDR < 0.01,]),
  rownames(fry.res.D5[fry.res.D5$FDR < 0.01,]),
  rownames(fry.res.D6[fry.res.D6$FDR < 0.01,])
  )
  )


hits.D3 <- topTags(qlf.D3, n = Inf)[[1]] %>% rownames_to_column() %>% filter(Symbol %in% hits)
hits.D3$timepoint <- "D3"
hits.D4 <- topTags(qlf.D4, n = Inf)[[1]] %>% rownames_to_column() %>% filter(Symbol %in% hits)
hits.D4$timepoint <- "D4"
hits.D5 <- topTags(qlf.D5, n = Inf)[[1]] %>% rownames_to_column() %>% filter(Symbol %in% hits)
hits.D5$timepoint <- "D5"
hits.D6 <- topTags(qlf.D6, n = Inf)[[1]] %>% rownames_to_column() %>% filter(Symbol %in% hits)
hits.D6$timepoint <- "D6"

hits.all <- rbind(hits.D3, hits.D4, hits.D5, hits.D6)
counts <- dge$counts
cpm <- edgeR::cpm(dge, normalized.lib.sizes = T, log = F)
rownames(cpm) <- dge$genes$Symbol

write.table(hits.all, file = file.path(results.dir, "32_58_D3_NEG_hits_all_timepoints.csv"), sep = ",")
```

## plot normalised counts per timepoint per guide of significant genes
```{r, eval = F}
pdf(file.path(plots.dir, "32_58_NEG_norm_counts_per_guide_timepoint.pdf"))
lapply(hits, function(x){
  dat <- as.data.frame(cpm) %>% 
    rownames_to_column() %>% 
    filter(grepl(as.character(x), rowname)) %>% 
    reshape2::melt() %>% 
    filter(grepl("NEG", variable)) %>% 
    mutate(timepoint = str_split(variable, "\\.", simplify = T)[,3])
  
    p <- ggplot(dat, aes(y = value, x = timepoint, color = rowname, group = 1)) + 
    geom_point() +
    geom_smooth(method = "loess") + 
    scale_y_continuous(trans = "log10") +
    facet_wrap(~rowname, scales = "free_y") + 
    labs(x = "Timepoint", y = "Log10 Counts Per Million") +
    theme_bw()
  return(p)
})
dev.off()
```

## plot log2FC per timepoint per guide of significant genes
```{r, eval = F}
pdf(file.path(plots.dir, "32_58_NEG_logFC_per_guide_timepoint.pdf"))
lapply(hits, function(x){
  dat <- as.data.frame(hits.all) %>% 
    filter(Symbol == as.character(x)) %>% 
    mutate(gene = paste(Symbol, Guide, sep = '.')) %>% 
    reshape2::melt()
  
    p <- ggplot(dat, aes(y = value, x = timepoint, color = rowname, group = 1)) + 
    geom_point() +
    geom_smooth(method = "loess") + 
    facet_wrap(~ gene, scales = "free_y") + 
    labs(x = "Timepoint", y = "Log2 FC vs corresponding timepoint LIB control") +
    theme_bw()
  return(p)
})
dev.off()
```

## plot pct abundance per timepoint per guide of significant genes
```{r, eval = F}
pdf(file.path(plots.dir, "32_58_NEG_pct_per_guide_timepoint.pdf"))
lapply(hits, function(x){
  dat <- as.data.frame(sweep(as.data.frame(cpm), 2, colSums(as.data.frame(cpm)), `/`) * 100) %>%
  rownames_to_column() %>% 
  filter(grepl(as.character(x), rowname)) %>% 
  reshape2::melt() %>% 
  filter(grepl("NEG", variable)) %>% 
  mutate(timepoint = str_split(variable, "\\.", simplify = T)[,3])
  
    p <- ggplot(dat, aes(y = value, x = timepoint, color = rowname, group = 1)) + 
    geom_point() +
    geom_smooth(method = "loess") + 
    facet_wrap(~rowname, scales = "free_y") + 
    labs(x = "Timepoint", y = "Percentage abundance") +
    theme_bw()
  return(p)
})
dev.off()

```

## Barcode plots
```{r}
barcodeplot(qlf.D3$table$logFC, index = genesymbollist$MED30, main = "Barcodeplot for MED30",
labels = c("Negative logFC", "Positive logFC"), quantile = c(-0.5, 0.5))
barcodeplot(qlf.D4$table$logFC, index = genesymbollist$MED30, main = "Barcodeplot for MED30",
labels = c("Negative logFC", "Positive logFC"), quantile = c(-0.5, 0.5))
barcodeplot(qlf.D5$table$logFC, index = genesymbollist$MED30, main = "Barcodeplot for MED30",
labels = c("Negative logFC", "Positive logFC"), quantile = c(-0.5, 0.5))
barcodeplot(qlf.D6$table$logFC, index = genesymbollist$MED30, main = "Barcodeplot for MED30",
labels = c("Negative logFC", "Positive logFC"), quantile = c(-0.5, 0.5))
```

## CPM X logFC PLOTS
```{r}
NEG.hits <- unique(hits.all$Symbol)
NEG.avgLogFC <- top4$table %>% select(Symbol, logFC, logCPM) %>% filter(Symbol %in% NEG.hits) %>% group_by(Symbol) %>% summarise(avgLogFC = mean(logFC))
NEG.avgLogCPM <- top4$table %>% select(Symbol, logFC, logCPM) %>% filter(Symbol %in% NEG.hits) %>% group_by(Symbol) %>% summarise(avgLogCPM = mean(logCPM))

dat <- camera.res.D3 %>% rownames_to_column(var = "Symbol") %>% filter(FDR < 0.1) 
dat <- left_join(dat, NEG.avgLogFC, by = "Symbol")
dat <- left_join(dat, NEG.avgLogCPM, by = "Symbol")

med.dat <- dat %>% filter(Symbol %in% c("MED15", "MED12"))
pol.dat <- dat[grep("POL", dat$Symbol),]

ggplot(dat, aes(x = avgLogFC, y = -log10(FDR)), alpha = 0.5) + 
  geom_point() +
  geom_point(data = med.dat, aes(x = avgLogFC, y = -log10(FDR), color = "red")) +
  scale_y_continuous(trans = "log10") + theme_bw() + 
  ggrepel::geom_text_repel(data = med.dat, aes(x = avgLogFC, y = -log10(FDR)),
    label=med.dat$Symbol) + xlim(c(0,3.5))

```

## Dotplots across timepoints
```{r}
# clone58.R2.D3.gene.mid$Timepoint = "58.R2.D3.MID"
# clone58.R2.D4.gene.mid$Timepoint = "58.R2.D4.MID"
# clone58.R2.D5.gene.mid$Timepoint = "58.R2.D5.MID"
# clone58.R2.D6.gene.mid$Timepoint = "58.R2.D6.MID"

# NEG samples
all.samps <- rbind(hits.D3, hits.D4, hits.D5, hits.D6)

all.samps <- as_tibble(all.samps)
all.samps.genes <- all.samps %>% filter(FDR < 0.05) %>% filter(logFC > 1) 
all.samps <- all.samps[which(all.samps$Symbol %in% all.samps.genes$Symbol),]
all.samps$Timepoint <- as.factor(all.samps$timepoint)
all.samps$Symbol <- as.factor(all.samps$Symbol)

# filter out max logFC value sgRNA per gene
all.samps.maxFC <- all.samps %>% group_by(Symbol, Timepoint) %>% filter(logFC == max(logFC))

# # MID samples
# all.samps.mid <- rbind(clone58.R2.D3.gene.mid, clone58.R2.D4.gene.mid, clone58.R2.D5.gene.mid, clone58.R2.D6.gene.mid)
# all.samps.mid <- as_tibble(all.samps.mid)
# all.samps.mid.genes <- all.samps.mid %>% filter(FDR < 0.05) %>% filter(Score > 1) 
# all.samps.mid <- all.samps.mid[which(all.samps.mid$id %in% all.samps.mid.genes$id),]
# all.samps.mid$Timepoint <- as.factor(all.samps.mid$Timepoint)
# all.samps.mid$id <- as.factor(all.samps.mid$id)

# setup color palette
myPalette <- colorRampPalette(brewer.pal(9, "Reds"))

# create plot
ggplot(all.samps.maxFC, aes(x = Timepoint, y = Symbol, size = -log10(FDR), col = logFC)) + 
  geom_point(stat = "identity", alpha = 0.8, shape = 16) +
  theme_bw() +
  scale_size_continuous(range = c(0.001, 6)) +
  labs(y = "Symbol", x = "Timepoint", title = "Clone 58 NEG screen hits") +
  theme(axis.text.x=element_text(angle = 0, size = 9), axis.text.y=element_text(size = 9)) + 
  scale_colour_gradientn(colours = myPalette(100), limits=c(1, 4), na.value = "grey90")
ggsave(file.path(plots.dir, "NEG_significant_genes_dotplot.pdf"), width = 5, height = 9)

# ggplot(all.samps.mid, aes(x = Timepoint, y = id, size = LogFDR, col = Score)) + 
#   geom_point(stat = "identity", alpha = 0.8, shape = 16) +
#   theme_bw() +
#   #scale_size_continuous(range = c(0.01, 8)) +
#   labs(y = "Gene", x = "Timepoint", title = "Clone 58 MID screen hits") +
#   theme(axis.text.x=element_text(angle = 0, size = 9), axis.text.y=element_text(size = 9)) +
#   scale_colour_gradientn(colours = myPalette(100), limits=c(1, 3), na.value = "grey90")
# ggsave("K562_Chromatin/58_chromatin/MID_significant_genes_dotplot.pdf", width = 4, height = 8)

```

## Mediator specific dotplots
```{r}
med.samps <- all.samps.maxFC[grep("^MED|CDK8|CCNC",all.samps.maxFC$Symbol, value = F),]
med.samps$Timepoint <- as.factor(med.samps$Timepoint)
med.samps$Symbol <- factor(med.samps$Symbol, levels = rev(c("MED14","MED6","MED8","MED11","MED17","MED18","MED20","MED1","MED4","MED9","MED19","MED21","MED31","MED15","MED27","MED29","MED30","MED16","MED23","MED24","MED25","CDK8","MED12","MED13")))


ggplot(med.samps, aes(x = Timepoint, y = Symbol, size = -log10(FDR), col = logFC)) + 
  geom_point(stat = "identity", alpha = 0.8, shape = 16) +
  theme_bw() +
  scale_size_continuous(range = c(0.01, 6)) +
  labs(y = "Gene", x = "Timepoint", title = "Clone 58 NEG Mediator hits") +
  theme(axis.text.x=element_text(angle = 0, size = 9), axis.text.y=element_text(size = 9)) + 
  scale_colour_gradientn(colours = myPalette(100), limits=c(1, 4), na.value = "grey90")
ggsave(file.path(plots.dir, "NEG_Mediator_genes_dotplot.pdf"), width = 4, height = 6)
```

## Pol II specific dotplots
```{r}
pol.samps <- all.samps.maxFC[grep("^POLR",all.samps.maxFC$Symbol, value = F),]
pol.samps$Timepoint <- as.factor(pol.samps$Timepoint)
pol.samps$Symbol <- factor(pol.samps$Symbol, 
                           levels = rev(c("POLR2A","POLR2B","POLR2C","POLR2D","POLR2E","POLR2F","POLR2G","POLR2H","POLR2L","POLR2K")))

ggplot(pol.samps, aes(x = Timepoint, y = Symbol, size = -log10(FDR), col = logFC)) + 
  geom_point(stat = "identity", alpha = 0.8, shape = 16) +
  theme_bw() +
  scale_size_continuous(range = c(0.01, 6)) +
  labs(y = "Gene", x = "Timepoint", title = "Clone 58 NEG Pol2 hits") +
  theme(axis.text.x=element_text(angle = 0, size = 9), axis.text.y=element_text(size = 9)) + 
  scale_colour_gradientn(colours = myPalette(100), limits=c(1, 4), na.value = "grey90")
ggsave(file.path(plots.dir, "NEG_Pol_genes_dotplot.pdf"), width = 4, height = 4)

```


## plot factor enrichment over time
```{r}
hits.dat <- all.samps.maxFC[which(all.samps.maxFC$Symbol %in% hits),]
hits.dat$Timepoint <- as.factor(hits.dat$Timepoint)
#hits.dat <- arrange(hits.dat, desc(logFC)) %>% filter(FDR < 0.1)

# define gene groups
hits.dat$Group <- "Other"
hits.dat$Group[grep("^MED|CDK8|CCNC",hits.dat$Symbol, value = F)] <- "Mediator"
hits.dat$Group[grep("^POLR",hits.dat$Symbol, value = F)] <- "Pol II"
hits.dat$Group[grep("EP300|BRD|SUPT|TAF|WDR|INTS|GTF",hits.dat$Symbol, value = F)] <- "Cofactor"

# smooth and line
hits.dat %>% ungroup() %>%
  dplyr::select(Symbol, logFC, Timepoint, Group) %>%
  reshape2::melt() %>% mutate(Group = factor(Group, levels = c("Pol II", "Mediator", "Cofactor", "Other"))) %>%
  ggplot(aes(x = Timepoint, y = value, group = Group, color = Group)) +
    geom_smooth() +
    geom_line(aes(x = Timepoint, y = value, group = Symbol), alpha = 0.2) +
    labs(x = "Timepoint", y = "logFC", title = "logFC by Timepoint") +
    theme_bw() + theme(legend.position = "none") +
    facet_wrap(~Group, nrow = 1)
ggsave("./edgeR/plots/group_enrichment_vs_time_smoothline.pdf", width = 6, height = 3)

# smooth only
hits.dat %>% ungroup() %>%
  dplyr::select(Symbol, logFC, Timepoint, Group) %>%
  reshape2::melt() %>% mutate(Group = factor(Group, levels = c("Pol II", "Mediator", "Cofactor", "Other"))) %>%
  ggplot(aes(x = Timepoint, y = value, group = Group, color = Group)) +
    geom_smooth() +
    labs(x = "Timepoint", y = "logFC", title = "logFC by Timepoint") +
    theme_bw() + theme(legend.position = "none") +
    facet_wrap(~Group, nrow = 1)
ggsave("./edgeR/plots/group_enrichment_vs_time_smooth.pdf", width = 6, height = 3)


# as boxplot
hits.dat %>% ungroup() %>%
  dplyr::select(Symbol, logFC, Timepoint, Group) %>%
  melt() %>%
  ggplot(aes(x = Timepoint, y = value, color = Timepoint)) +
    geom_boxplot() +
    labs(x = "Timepoint", y = "logFC", title = "logFC by Timepoint") +
    theme_bw() + theme(legend.position = "none") +facet_wrap(~Group, nrow = 1)
ggsave("./edgeR/plots/group_enrichment_vs_time_box.pdf", width = 6, height = 3)

```

## SAGA specific dotplots - only SUPT20H and TAF10 hit
```{r}
# saga.genes <- c("SF3B3", "SF3B5", "TRAAP", "KAT2A", "KAT2B","SGF29", "TADA2B", "TADA3", "TBP", "SUPT3H", "SUPT7L", "SUPT20H", "TADA1", "TAF5L", "TAF6L",
#                 "TAF9B","TAF10", "TAF12", "ATXN7", "ATXN7L3", "ENY2", "USP22")
# saga.samps <- all.samps.maxFC %>% filter(Symbol %in% saga.genes)
# saga.samps$Timepoint <- as.factor(saga.samps$Timepoint)
# saga.samps$Symbol <- factor(saga.samps$Symbol, 
#                            levels = rev(c()))
# 
# ggplot(pol.samps, aes(x = Timepoint, y = Symbol, size = -log10(FDR), col = logFC)) + 
#   geom_point(stat = "identity", alpha = 0.8, shape = 16) +
#   theme_bw() +
#   #scale_size_continuous(range = c(0.01, 8)) +
#   labs(y = "Gene", x = "Timepoint", title = "Clone 58 NEG screen hits") +
#   theme(axis.text.x=element_text(angle = 0, size = 9), axis.text.y=element_text(size = 9)) + 
#   scale_colour_gradientn(colours = myPalette(100), limits=c(1, 4), na.value = "grey90")
# ggsave("K562_Chromatin/58_chromatin/NEG_significant_genes_dotplot.pdf", width = 4, height = 8)

```

## Dotplots per day - needs fixing
```{r}
dat <- as.tibble(all.samps.maxFC)
dat$LogFDR <- -log10(dat$FDR)
dat$RandomIndex <- sample(1:nrow(dat), nrow(dat))
alphabetical <- dat[order(dat$Symbol),]
dat$alphabetical <- seq(1,nrow(dat))
dat <- dat[order(-dat$logFC), ]
pos <- dat[dat$logFC>0, ]
pos <- pos %>% mutate(color = if_else(pos$LogFDR>1.3,true = "Sig", false = "NotSig"))
neg <- dat[dat$logFC<0, ]
dat <- dat %>% mutate(color = if_else(dat$LogFDR>1.3,true = "Sig", false = "NotSig"))
dat$label <- dat$Symbol
dat$label <- if_else(dat$LogFDR>1.3, dat$label, "")
ixlabel <- which(dat$LogFDR>1.3)
dat$label[ixlabel] <- dat$Symbol[ixlabel] 
ggplot(dat) + 
  geom_point(aes(x = alphabetical, y = LogFDR, col = color, label = Symbol)) + 
  theme_bw() + 
  scale_y_continuous(trans = "log1p") + 
  ylim(c(0,4)) +
  geom_hline(yintercept = 1.3, color = "grey40") +
  scale_size_continuous(range = c(-6, 6)) +
  scale_colour_manual(values = c("grey", "firebrick")) +
  facet_wrap(~ Timepoint, scales = "free_x")

ggplot(pos) + 
  geom_point(aes(x = RandomIndex, y = LogFDR, col = color, label = id)) + 
  theme_bw() + 
  scale_y_continuous(trans = "log1p") + 
  ylim(c(0,4)) +
  geom_hline(yintercept = 1.3, color = "grey40") +
  scale_size_continuous(range = c(-6, 6)) +
  scale_colour_manual(values = c("grey", "firebrick"))


  
ScatterView(dat, x = "RandomIndex", y = "Score", label = "id", size = "LogFDR", 
                 y_cut = 1.3, groups = "top", toplabels = pos$id[1:10], force = 2, , display_cut = T) + ylim(c(0,4)) 
p
```

#----------- Paper figures
## import hgnc gene complex annotations and replot
```{r}
hgnc.anno <- read.csv("./hgnc-symbol-check.csv")
hgnc.anno <- as.tibble(hgnc.anno)

hits.dat <- all.samps.maxFC[which(all.samps.maxFC$Symbol %in% hits),]
hits.dat$Timepoint <- as.factor(hits.dat$Timepoint)
#hits.dat <- arrange(hits.dat, desc(logFC)) %>% filter(FDR < 0.1)

# match gene groups
which(hgnc.anno$Approved.symbol %in% hits.dat$Symbol)
group.info <- lapply(hits.dat$Symbol, function(x){
  group <- hgnc.anno[which(hgnc.anno$Input == x, arr.ind = TRUE),]
  group <- group[1,7]
  return(group)
})

hits.dat$Group <- unlist(group.info)
table(hits.dat$Group)

# smooth and line
hits.dat %>% ungroup() %>%
  dplyr::select(Symbol, logFC, Timepoint, Group) %>%
  reshape2::melt() %>% mutate(Group = factor(Group, levels = c("RNA Pol II", "Mediator", "Core transcription", "Transcriptional initiation pausing elongation", "Chromatin modifying", "TF", "Structural", "Other"))) %>% filter(Group != "Other" & Group != "Structural") %>%
  ggplot(aes(x = Timepoint, y = value, group = Group, color = Group)) +
    geom_smooth() +
    geom_line(aes(x = Timepoint, y = value, group = Symbol), alpha = 0.2) +
    labs(x = "Timepoint", y = "logFC", title = "logFC by Timepoint") +
    theme_bw() + theme(legend.position = "none") +
    facet_wrap(~Group, nrow = 1)
ggsave("./edgeR/plots/group_enrichment_vs_time_smoothline_regroup.pdf", width = 6, height = 3)

# smooth only
hits.dat %>% ungroup() %>%
  dplyr::select(Symbol, logFC, Timepoint, Group) %>%
  reshape2::melt() %>% mutate(Group = factor(Group, levels = c("RNA Pol II", "Core transcription", "Mediator", "Chromatin modifying", "Transcriptional initiation pausing elongation", "TF", "Structural", "Other"))) %>%
  ggplot(aes(x = Timepoint, y = value, group = Group, color = Group)) +
    geom_smooth() +
    labs(x = "Timepoint", y = "logFC", title = "logFC by Timepoint") +
    theme_bw() + theme(legend.position = "none") +
    facet_wrap(~Group, nrow = 1)
ggsave("./edgeR/plots/group_enrichment_vs_time_smooth_regroup.pdf", width = 6, height = 3)


# as boxplot
hits.dat %>% ungroup() %>%
  dplyr::select(Symbol, logFC, Timepoint, Group) %>%
  melt() %>%
  ggplot(aes(x = Timepoint, y = value, color = Timepoint)) +
    geom_boxplot() +
    labs(x = "Timepoint", y = "logFC", title = "logFC by Timepoint") +
    theme_bw() + theme(legend.position = "none") +facet_wrap(~Group, nrow = 1)
ggsave("./edgeR/plots/group_enrichment_vs_time_box_regroup.pdf", width = 6, height = 3)


```

## Dotplots across timepoints - alternate
```{r}

# NEG samples
all.dat <- rbind(hits.D3, hits.D4, hits.D5, hits.D6)
all.samps <- all.dat %>% filter(FDR < 0.05 | logFC > 1) # 
#all.samps <- all.samps %>% filter(Symbol %in% hits) 
all.samps <- all.samps %>% filter(Symbol != "safe") %>% filter(Symbol != "none")
genes.to.plot <- unique(all.samps$Symbol)

all.samps <- all.dat %>% filter(Symbol %in% genes.to.plot)
all.samps$Timepoint <- as.factor(all.samps$timepoint)
all.samps$Symbol <- as.factor(all.samps$Symbol)

group.info <- lapply(all.samps$Symbol, function(x){
  group <- hgnc.anno[which(hgnc.anno$Input == x, arr.ind = TRUE),]
  group <- group[1,7]
  return(group)
})

all.samps$Group <- unlist(group.info)
table(all.samps$Group)

# filter out max logFC value sgRNA per gene per timepoint
all.samps.maxFC <- all.samps %>% 
  group_by(Symbol, Timepoint) %>% 
  filter(logFC == max(logFC)) %>%
  mutate(Group = factor(Group, levels = c("RNA Pol II", "Core transcription", "Mediator", "Chromatin modifying", "Transcriptional initiation pausing elongation", "TF", "Structural", "Other"))) %>% filter(Group != "Other" & Group != "Structural") 

# reset gene groups to plot based on gene family
all.samps.maxFC <- all.samps.maxFC[order(all.samps.maxFC$Group),]
all.samps.maxFC$Symbol <- factor(all.samps.maxFC$Symbol, levels = rev(unique(all.samps.maxFC$Symbol)))

# setup color palette
myPalette <- colorRampPalette(brewer.pal(100, "Reds"))

# create plot
library(forcats)

p <- ggplot(all.samps.maxFC, aes(x = Timepoint, y = Symbol, size = -log10(FDR), col = logFC)) + 
  geom_point(stat = "identity", alpha = 0.8, shape = 16) +
  theme_bw() +
  scale_size_continuous(range = c(0.001, 6)) +
  labs(y = "Symbol", x = "Timepoint", title = "Clone 58 NEG screen hits") +
  theme(axis.text.x=element_text(angle = 0, size = 9), axis.text.y=element_text(size = 9)) +
  theme_classic() +
  scale_colour_gradientn(colours = myPalette(100), limits=c(1, 4), na.value = "grey80")
ggsave(file.path(plots.dir, "NEG_significant_genes_dotplot.pdf"), width = 5, height = 9)

library(cowplot)

# Create the annotation bar (side bar)
annotation <- ggplot(all.samps.maxFC, aes(x = 1, y = Symbol, fill = Group)) +
  geom_tile() +
  theme_void() + 
  theme(legend.position = "none")

# Combine the dot plot and annotation bar
combined_plot <- p + annotation + plot_layout(ncol = 2, widths = c(4, 1))

combined_plot

# ggplot(all.samps.mid, aes(x = Timepoint, y = id, size = LogFDR, col = Score)) + 
#   geom_point(stat = "identity", alpha = 0.8, shape = 16) +
#   theme_bw() +
#   #scale_size_continuous(range = c(0.01, 8)) +
#   labs(y = "Gene", x = "Timepoint", title = "Clone 58 MID screen hits") +
#   theme(axis.text.x=element_text(angle = 0, size = 9), axis.text.y=element_text(size = 9)) +
#   scale_colour_gradientn(colours = myPalette(100), limits=c(1, 3), na.value = "grey90")
# ggsave("K562_Chromatin/58_chromatin/MID_significant_genes_dotplot.pdf", width = 4, height = 8)

```

#-----------

#------------------------------------------------
# GSEA / GO analysis
#------------------------------------------------

## setup enrichR
```{r}
library(enrichR)

websiteLive <- getOption("enrichR.live")
if (websiteLive) {
    listEnrichrSites()
    setEnrichrSite("Enrichr") # Human genes   
}

dbs <- listEnrichrDbs()
enriched <- enrichr(c("Runx1", "Gfi1", "Gfi1b", "Spi1", "Gata1", "Kdr"), dbs)

dbs <- c("GO_Molecular_Function_2015", "GO_Cellular_Component_2015", "GO_Biological_Process_2015")


plotEnrich(enriched[[3]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value")

all.samps.maxFC

geneList <- clone58.R2.D6.gene$FDR
names(geneList) <- clone58.R2.D6.gene$id
enrich <- EnrichAnalyzer(geneList = geneList[geneList<0.05], 
                        method = "HGT", type = "GOCC", filter = F)
EnrichedView(enrich, mode = 1, top = 20)

#enrichPro = EnrichAnalyzer(geneList[geneList>1], type = "CORUM+CPX", filter = T)
#EnrichedView(enrichPro, top = 30)
```

#------------------------------------------------
# MID sample analysis - Combined clones
#------------------------------------------------

