---
title: "K562 MYC-mNG Clone 58 Drug Screen = Chromatin library - IBET inhibitor"
output: html_document
---

#--------------------------------------------
# K562 MYC-mNG Clone 58 Drug Screen = Chromatin library - IBET inhibitor
#--------------------------------------------
K562 MYC-mNeonGreen cells were transduced with the human chromatin focussed sgRNA library. 
Cells were then treated with 2.5uM IBET-151
GFP positive cells that maintained MYC expression in the presence of IBET were enriched by FACS at 24h and 48h post drug addition and sequenced --> SORT fraction (positive selection arm)
Timepoint matched LIB control fractions were taken at each timepoint. 
Two replicates were performed, one each for both clone 3 and clone 5. 

BARtab v1.4 was used to extract and count sgRNAs from raw sequencing reads
Resulting counts tables are imported here for CRISPR screen analysis using edgeR.

# Setup

## import libraries
```{r, message=FALSE, warning=FALSE}
library(edgeR)
library(bartools)
library(tidyverse)
library(ggpointdensity)
library(viridis)
#library(MAGeCKFlute)
library(RColorBrewer)
library(ggvenn)
library(enrichR)
```

## setup output dirs
```{r, message=FALSE, warning=FALSE}
dir.create("edgeR", showWarnings = F)

results.dir <- file.path(getwd(), "edgeR/results")
dir.create(results.dir, showWarnings = F)

plots.dir <- file.path(getwd(), "edgeR/plots")
dir.create(plots.dir, showWarnings = F)
```

#------------------------------------------------
# Setup DGEList object
#------------------------------------------------

## import counts files
```{r}
raw.counts <- read.delim("./K562_IBET_chromatin_raw_counts.txt", header = T, sep = "\t", row.names = 1)
raw.counts <- raw.counts[,grep("BET", colnames(raw.counts))]
colnames(raw.counts) <- c("CHROM.24h.IBET.SORT.R1", "CHROM.48h.IBET.SORT.R1", "CHROM.48h.IBET.LIB.R1", "CHROM.24h.IBET.LIB.R2", "CHROM.24h.IBET.LIB.R1", "CHROM.48h.IBET.SORT.R2", "CHROM.24h.IBET.SORT.R2", "CHROM.48h.IBET.LIB.R2")
```

## generate metadata
```{r}
metadata <- as.data.frame(str_split(colnames(raw.counts), pattern = "\\.", simplify = T), row.names = colnames(raw.counts))
colnames(metadata) <- c("Library", "Timepoint", "Treatment", "Group", "Replicate")
metadata$Contrasts <- paste(metadata$Group, metadata$Timepoint, sep = ".")

# fix gene names
genes <- as.data.frame(str_split(rownames(raw.counts), "_", simplify = T))
sgRNAs <- as.data.frame(str_split(genes[,3], "\\.", simplify = T))
genes$V4 <- sgRNAs$V2
genes$V3 <- sgRNAs$V1
colnames(genes) <- c("Symbol", "ENSEMBL", "ENTREZ", "Guide")
```

## setup DGEList
```{r}
dge <- DGEList(counts = as.matrix(raw.counts), samples = metadata, group = metadata$Group, genes = genes, remove.zeros = T)
dim(dge)
```

## Filter sgRNAs
```{r}
table(dge$samples$group)
# Filter sgRNAs and samples with low counts Need a CPM greater than 5 in 15 or more samples
# to keep sgRNAs
selr = rowSums(dge$counts) > 10
table(selr)
# Need at least 100,000 reads to keep a given sample
selc = colSums(dge$counts) >= 1e+05
table(selc)
dge = dge[selr, selc]
dim(dge)
saveRDS(dge, file.path(results.dir, "./CHROM_IBET_DGE.rds"))
```

## Plot counts per sample & guide
```{r}
# Set up drug treatment colours
cols = as.numeric(dge$samples$group) + 1
# Plot number of sgRNAs that could be matched per sample and total for each sgRNA across
# all samples

barplot(colSums(dge$counts), las = 2, main = "Counts per index", col = cols, cex.names = 0.5,
cex.axis = 0.8)
barplot(rowSums(dge$counts), las = 2, main = "Counts per sgRNA", cex.names = 0.5, cex.axis = 0.8)

pdf(file.path(plots.dir,"./CHROM_IBET_counts_per_index.pdf"), useDingbats = F)
barplot(colSums(dge$counts), las = 2, main = "Counts per index", col = cols, cex.names = 0.5, cex.axis = 0.8)
barplot(rowSums(dge$counts), las = 2, main = "Counts per sgRNA", cex.names = 0.5, cex.axis = 0.8)
dev.off()
```
#------------------------------------------------
# Differential abundance analysis: SORT vs LIB
#------------------------------------------------

```{r}
dge = calcNormFactors(dge, method = "TMM")

# Make an MDS plot to visualise relationships between replicate samples Set up infection #
# colours
cols2 = as.numeric(factor(dge$samples$group)) + 1
plotMDS(removeBatchEffect(dge, dge$samples$Replicate), col = cols, main = "MDS Plot - limma::removeBatchEffect")
plotMDS(dge, col = cols2, main = "MDS Plot - no batch correction")

pdf(file.path(plots.dir, "CHROM_IBET_MDS.pdf"), useDingbats = F)
plotMDS(removeBatchEffect(dge, dge$samples$Replicate), col = cols, main = "MDS Plot - limma::removeBatchEffect")
plotMDS(dge, col = cols2, main = "MDS Plot - no batch correction")
dev.off()
```
## Set up design matrix for GLM
```{r}
# Set up design matrix for GLM
group = as.factor(dge$samples$group)
replicate = as.factor(dge$samples$Replicate)
timepoint = as.factor(dge$samples$Timepoint)
contrasts = as.factor(dge$samples$Contrasts)

des = model.matrix(~0 + contrasts)
#colnames(des) <- levels(sample)
colnames(des) <- c("LIB.24h", "LIB.48h", "SORT.24h", "SORT.48h")
des

```

## Estimate dispersions
```{r}
xglm = estimateDisp(dge, des, tagwise = T)
sqrt(xglm$common.disp)
```

## Plot BCVs versus abundance
```{r}
plotBCV(xglm, main = "BCV Plot")

# save BCV plot
pdf(file.path(plots.dir,"CHROM_BETi_BCV.pdf"), useDingbats = F)
plotBCV(xglm, main = "BCV Plot")
dev.off()
```

## Fit negative bionomial GLM
```{r}
fit = glmFit(xglm, des)
fit.qlf = glmQLFit(xglm, des)

# define contrasts of interest
# separate timepoints
my.contrasts <- makeContrasts(
  SORT.D1 = SORT.24h-LIB.24h,
  SORT.D2 = SORT.48h-LIB.48h,
  levels = des)

# Carry out LRT test
lrt.D1 = glmLRT(fit, contrast = my.contrasts[,"SORT.D1"])
lrt.D2 = glmLRT(fit, contrast = my.contrasts[,"SORT.D2"])

write.csv(topTags(lrt.D1, n = Inf), file = file.path(results.dir, "CHROM_IBET_D1_LRT_toptags.csv"))
write.csv(topTags(lrt.D2, n = Inf), file = file.path(results.dir, "CHROM_IBET_D2_LRT_toptags.csv"))

# Carry out QLF test
qlf.D1 = glmQLFTest(fit.qlf, contrast = my.contrasts[,"SORT.D1"])
qlf.D2 = glmQLFTest(fit.qlf, contrast = my.contrasts[,"SORT.D2"])

write.csv(topTags(qlf.D1, n = Inf), file = file.path(results.dir, "CHROM_IBET_D1_qlf_toptags.csv"))
write.csv(topTags(qlf.D2, n = Inf), file = file.path(results.dir, "CHROM_IBET_D2_qlf_toptags.csv"))
```

## Plot smear for LO samples
```{r}
# SORT fraction D1
thresh = 0.1
lfc = 1
top4 = topTags(qlf.D1, n = Inf)
top4ids = rownames(top4$table[top4$table$FDR < thresh & abs(top4$table$logFC) >= lfc,])

plotSmear(qlf.D1, de.tags = top4ids, pch = 20, cex = 0.6, main = "IBET 24h CHROM SORT: logFC vs logCPM")
abline(h = c(-1, 0, 1), col = c("dodgerblue", "yellow", "dodgerblue"), lty = 2)

pdf(file.path(plots.dir,"CHROM_IBET_smear_D1.pdf"), useDingbats = F)
plotSmear(qlf.D1, de.tags = top4ids, pch = 20, cex = 0.6, main = "IBET 24h CHROM SORT: logFC vs logCPM")
abline(h = c(-1, 0, 1), col = c("dodgerblue", "yellow", "dodgerblue"), lty = 2)
dev.off()

# SORT fraction D2
thresh = 0.1
lfc = 1
top4 = topTags(qlf.D2, n = Inf)
top4ids = rownames(top4$table[top4$table$FDR < thresh & abs(top4$table$logFC) >= lfc,])

plotSmear(qlf.D2, de.tags = top4ids, pch = 20, cex = 0.6, main = "IBET 48h CHROM SORT: logFC vs logCPM")
abline(h = c(-1, 0, 1), col = c("dodgerblue", "yellow", "dodgerblue"), lty = 2)

pdf(file.path(plots.dir,"CHROM_IBET_smear_D2.pdf"), useDingbats = F)
plotSmear(qlf.D2, de.tags = top4ids, pch = 20, cex = 0.6, main = "IBET 48h CHROM SORT: logFC vs logCPM")
abline(h = c(-1, 0, 1), col = c("dodgerblue", "yellow", "dodgerblue"), lty = 2)
dev.off()
```


#------------------------------------------------
# Gene set testing:
#------------------------------------------------

## setup gene list for camera
```{r}
genesymbols = dge$genes[,1]
genesymbollist = list()
unq = unique(genesymbols)
unq = unq[!is.na(unq)]
for (i in unq) {
  sel = genesymbols == i & !is.na(genesymbols)
  if (sum(sel) > 3){
    genesymbollist[[i]] = which(sel)
  }
}

write(unq, file = file.path(results.dir, "CHROM_lib_genes.txt"))
```

## Run camera for all genes
```{r}
camera.res.D1 = camera(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D1"])
camera.res.D1 %>% filter(Direction == "Up") %>% arrange(FDR)
camera.res.D1[camera.res.D1$FDR < 0.1,]

camera.res.D2 = camera(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D2"])
camera.res.D2 %>% filter(Direction == "Up") %>% arrange(FDR)
camera.res.D2[camera.res.D2$FDR < 0.1,]

```

## Run roast for all genes
```{r}
roast.res.D1 = roast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D1"])
roast.res.D1 %>% filter(Direction == "Up") %>% arrange(FDR)
roast.res.D1[roast.res.D1$FDR < 0.1,]

roast.res.D2 = roast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D2"])
roast.res.D2 %>% filter(Direction == "Up") %>% arrange(FDR)
roast.res.D2[roast.res.D2$FDR < 0.1,]

```

## Run mroast for all genes
```{r}
mroast.res.D1 = mroast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D1"])
mroast.res.D1 %>% filter(Direction == "Up") %>% arrange(FDR)
mroast.res.D1[mroast.res.D1$FDR < 0.1,]

mroast.res.D2 = mroast(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D2"])
mroast.res.D2 %>% filter(Direction == "Up") %>% arrange(FDR)
mroast.res.D2[mroast.res.D2$FDR < 0.1,]

```

## Run fry for all genes
```{r}
fry.res.D1 = fry(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D1"])
fry.res.D1 %>% filter(Direction == "Up") %>% arrange(FDR)
fry.res.D1[fry.res.D1$FDR < 0.1,]

fry.res.D2 = fry(xglm, index = genesymbollist, des, contrast = my.contrasts[,"SORT.D2"])
fry.res.D2 %>% filter(Direction == "Up") %>% arrange(FDR)
fry.res.D2[fry.res.D2$FDR < 0.1,]

```

## write gene set results to file
```{r}
# camera
write.table(camera.res.D1, file = file.path(results.dir, "IBET_CHROM_D1_camera_res.csv"), sep = ",")
write.table(camera.res.D2, file = file.path(results.dir, "IBET_CHROM_D2_camera_res.csv"), sep = ",")

# roast
write.table(roast.res.D1, file = file.path(results.dir, "IBET_CHROM_D1_roast_res.csv"), sep = ",")
write.table(roast.res.D2, file = file.path(results.dir, "IBET_CHROM_D2_roast_res.csv"), sep = ",")

# mroast
write.table(mroast.res.D1, file = file.path(results.dir, "IBET_CHROM_D1_mroast_res.csv"), sep = ",")
write.table(mroast.res.D2, file = file.path(results.dir, "IBET_CHROM_D2_mroast_res.csv"), sep = ",")

# fry
write.table(fry.res.D1, file = file.path(results.dir, "IBET_CHROM_D1_fry_res.csv"), sep = ",")
write.table(fry.res.D2, file = file.path(results.dir, "IBET_CHROM_D2_fry_res.csv"), sep = ",")

```

## Gene set enrichment analysis - enrichR
```{r}
library(enrichR)
dbs <- enrichR::listEnrichrDbs()
dbs.to.use <- c("KEGG_2021_Human", 
                "GO_Biological_Process_2023", 
                "GO_Cellular_Component_2023", 
                "GO_Molecular_Function_2023", 
                "MSigDB_Hallmark_2020",
                "GeneSigDB", 
                "CORUM")
sig.genes.24h <- rownames(camera.res.D1[camera.res.D1$FDR < 0.1 & camera.res.D1$Direction == "Up",])
sig.genes.48h <- rownames(camera.res.D2[camera.res.D2$FDR < 0.1 & camera.res.D2$Direction == "Up",])

pdf(file.path(plots.dir,"CHROM_IBET_enrichR_D1.pdf"), useDingbats = F)
res <- enrichr(genes = sig.genes.24h, databases = dbs.to.use)
lapply(dbs.to.use, function(x){
  enrichR::plotEnrich(res[[as.character(x)]], title = as.character(x))
})
dev.off()

pdf(file.path(plots.dir,"CHROM_IBET_enrichR_D2.pdf"), useDingbats = F)
res <- enrichr(genes = sig.genes.48h, databases = dbs.to.use)
lapply(dbs.to.use, function(x){
  enrichR::plotEnrich(res[[as.character(x)]], title = as.character(x))
})
dev.off()

```

#------------------------------------------------
# PLOTS
#------------------------------------------------

## get sigificant genes from camera roast mroast and fry analysis
```{r}
hits <- unique(
  c(
  rownames(camera.res.D1[camera.res.D1$FDR < 0.01,]),
  rownames(camera.res.D2[camera.res.D2$FDR < 0.01,]),
  
  rownames(roast.res.D1[roast.res.D1$FDR < 0.01,]),
  rownames(roast.res.D2[roast.res.D2$FDR < 0.01,]),
  
  rownames(mroast.res.D1[mroast.res.D1$FDR < 0.01,]),
  rownames(mroast.res.D2[mroast.res.D2$FDR < 0.01,]),
  
  rownames(fry.res.D1[fry.res.D1$FDR < 0.01,]),
  rownames(fry.res.D2[fry.res.D2$FDR < 0.01,])
  )
  )


hits.D1 <- topTags(qlf.D1, n = Inf)[[1]] %>% rownames_to_column() %>% filter(Symbol %in% hits)
hits.D1$timepoint <- "D1"
hits.D2 <- topTags(qlf.D2, n = Inf)[[1]] %>% rownames_to_column() %>% filter(Symbol %in% hits)
hits.D2$timepoint <- "D2"

hits.all <- rbind(hits.D1, hits.D2)
counts <- dge$counts
cpm <- edgeR::cpm(dge, normalized.lib.sizes = T, log = F)
rownames(cpm) <- dge$genes$Symbol

write.table(hits.all, file = file.path(results.dir, "CHROM_IBET_hits_all_timepoints.csv"), sep = ",")
```

## plot normalised counts per timepoint per guide of significant genes
```{r}
pdf(file.path(plots.dir, "CHROM_IBET_norm_counts_per_guide_timepoint.pdf"))
lapply(hits, function(x){
  dat <- as.data.frame(cpm) %>%
    rownames_to_column() %>% 
    filter(grepl(as.character(x), rowname)) %>% 
    reshape2::melt() %>% 
    mutate(timepoint = str_split(variable, "\\.", simplify = T)[,2])
  
    p <- ggplot(dat, aes(y = value, x = timepoint, color = rowname, group = 1)) + 
    geom_point() +
    geom_smooth(method = "loess") + 
    scale_y_continuous(trans = "log10") +
    facet_wrap(~rowname, scales = "free_y") + 
    labs(x = "Timepoint", y = "Log10 Counts Per Million") +
    theme_bw()
  return(p)
})
dev.off()
```

## plot log2FC per timepoint per guide of significant genes
```{r}
pdf(file.path(plots.dir, "CHROM_IBET_logFC_per_guide_timepoint.pdf"))
lapply(hits, function(x){
  dat <- as.data.frame(hits.all) %>% 
    filter(Symbol == as.character(x)) %>% 
    mutate(gene = paste(Symbol, Guide, sep = '.')) %>% 
    reshape2::melt()
  
    p <- ggplot(dat, aes(y = value, x = timepoint, color = rowname, group = 1)) + 
    geom_point() +
    geom_smooth(method = "loess") + 
    facet_wrap(~ gene, scales = "free_y") + 
    labs(x = "Timepoint", y = "Log2 FC vs corresponding timepoint LIB control") +
    theme_bw()
  return(p)
})
dev.off()
```

## plot pct abundance per timepoint per guide of significant genes
```{r}
pdf(file.path(plots.dir, "CHROM_IBET_pct_per_guide_timepoint.pdf"))
lapply(hits, function(x){
  dat <- as.data.frame(sweep(as.data.frame(cpm), 2, colSums(as.data.frame(cpm)), `/`) * 100) %>%
  rownames_to_column() %>% 
  filter(grepl(as.character(x), rowname)) %>% 
  reshape2::melt() %>% 
  mutate(timepoint = str_split(variable, "\\.", simplify = T)[,2])
  
    p <- ggplot(dat, aes(y = value, x = timepoint, color = rowname, group = 1)) + 
    geom_point() +
    geom_smooth(method = "loess") + 
    facet_wrap(~rowname, scales = "free_y") + 
    labs(x = "Timepoint", y = "Percentage abundance") +
    theme_bw()
  return(p)
})
dev.off()

```

## Dotplots - IBET D1 - setup
```{r}
# merge data
IBET.camera.D1 <- merge(camera.res.D1, roast.res.D1, by=0, all.x=TRUE)

# convert FDR to -log10(FDR)
IBET.camera.D1$LogFDR = -log10(IBET.camera.D1$FDR.x)

# setup plot data
dat <- IBET.camera.D1
set.seed(103)

# give each gene a random position on x axis
dat$RandomIndex <- sample(1:nrow(dat), nrow(dat)) 

# rank genes alphabetically
alphabetical <- IBET.camera.D1[order(IBET.camera.D1$Row.names),]
dat$alphabetical <- seq(1,nrow(dat))

# order by significance
dat <- dat[order(-dat$LogFDR), ]

# filter positive enrichments
pos <- dat[dat$Direction.x == "Up", ]
pos <- pos %>% mutate(color = if_else(pos$LogFDR>2, true = "Sig", false = "NotSig"))

# setup significance grouping variables
# significant hits have -log10pval > 2 (padj < 0.01) and proportion guides positively enriched (PropUp) > 0.4
dat <- pos %>% mutate(color = if_else(pos$LogFDR>2,true = "Sig", false = "NotSig"))
dat$color[which(dat$PropUp < 0.4)] <- "NotSig"
dat$label <- as.character(dat$Row.names)
dat$label <- ifelse(dat$LogFDR>2 & dat$PropUp > 0.4 | dat$Row.names == "SF3B3" | dat$Row.names == "SF3B1", dat$label, "")
ixlabel <- which(dat$LogFDR>2)
dat$label[ixlabel] <- dat$Row.names[ixlabel]
d1.pos.dat.LOW <- pos
```

## Dotplots - LO reactivation D1 - plotting
```{r}
ggplot(dat, aes(RandomIndex, LogFDR, label = label, size = PropUp, col = color)) + 
  geom_point(alpha = 0.4) +
  scale_y_continuous(trans = "log1p") +
  scale_colour_manual(values = c("grey60", "green3")) +
  scale_size_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  theme_bw() +
  #ylim(c(0,3)) +
  geom_hline(yintercept = 2, color = "black", linetype = 3) +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 1000)
ggsave(filename = "edgeR/plots/CHROM_IBET_dotplots_SORT_D1_propup.pdf", width = 4, height = 4)

# alphabetical
ggplot(dat, aes(alphabetical, LogFDR, label = label, size = PropUp, col = color)) + 
  geom_point(alpha = 0.4) +
  scale_y_continuous(trans = "log1p") +
  scale_colour_manual(values = c("grey60", "green3")) +
  scale_size_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  theme_bw() +
  #ylim(c(0,3)) +
  geom_hline(yintercept = 2, color = "black", linetype = 3) +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 1000)
ggsave(filename = "edgeR/plots/CHROM_IBET_dotplots_SORT_D1_propup_alphabetical.pdf", width = 4, height = 4)

```

## Dotplots - IBET D2 - setup
```{r}
# merge data
IBET.camera.D2 <- merge(camera.res.D2, roast.res.D2, by=0, all.x=TRUE)

# convert FDR to -log10(FDR)
IBET.camera.D2$LogFDR = -log10(IBET.camera.D2$FDR.x)

# setup plot data
dat <- IBET.camera.D2
set.seed(103)

# give each gene a random position on x axis
dat$RandomIndex <- sample(1:nrow(dat), nrow(dat)) 

# rank genes alphabetically
alphabetical <- IBET.camera.D2[order(IBET.camera.D2$Row.names),]
dat$alphabetical <- seq(1,nrow(dat))

# order by significance
dat <- dat[order(-dat$LogFDR), ]

# filter positive enrichments
pos <- dat[dat$Direction.x == "Up", ]
pos <- pos %>% mutate(color = if_else(pos$LogFDR>2, true = "Sig", false = "NotSig"))

# setup significance grouping variables
# significant hits have -log10pval > 2 (padj < 0.01) and proportion guides positively enriched (PropUp) > 0.4
dat <- pos %>% mutate(color = if_else(pos$LogFDR>2,true = "Sig", false = "NotSig"))
dat$color[which(dat$PropUp < 0.4)] <- "NotSig"
dat$label <- as.character(dat$Row.names)
dat$label <- ifelse(dat$LogFDR>2 & dat$PropUp > 0.4 | dat$Row.names == "SF3B3" | dat$Row.names == "SF3B1", dat$label, "")
ixlabel <- which(dat$LogFDR>2)
dat$label[ixlabel] <- dat$Row.names[ixlabel]
D2.pos.dat.LOW <- pos
```

## Dotplots - LO reactivation D2 - plotting
```{r}
ggplot(dat, aes(RandomIndex, LogFDR, label = label, size = PropUp, col = color)) + 
  geom_point(alpha = 0.4) +
  scale_y_continuous(trans = "log1p") +
  scale_colour_manual(values = c("grey60", "green3")) +
  scale_size_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  theme_bw() +
  #ylim(c(0,3)) +
  geom_hline(yintercept = 2, color = "black", linetype = 3) +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 1000)
ggsave(filename = "edgeR/plots/CHROM_IBET_dotplots_SORT_D2_propup.pdf", width = 4, height = 4)

# alphabetical
ggplot(dat, aes(alphabetical, LogFDR, label = label, size = PropUp, col = color)) + 
  geom_point(alpha = 0.4) +
  scale_y_continuous(trans = "log1p") +
  scale_colour_manual(values = c("grey60", "green3")) +
  scale_size_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  theme_bw() +
  #ylim(c(0,3)) +
  geom_hline(yintercept = 2, color = "black", linetype = 3) +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 1000)
ggsave(filename = "edgeR/plots/CHROM_IBET_dotplots_SORT_D2_propup_alphabetical.pdf", width = 4, height = 4)

```

## combined timepoint dotplots
```{r}
# merge camera and roast results
IBET.24h.camera <- merge(camera.res.D1, roast.res.D1, by = 0, all = T)
IBET.24h.camera <- IBET.24h.camera %>% filter(Direction.x == "Up")
IBET.48h.camera <- merge(camera.res.D2, roast.res.D2, by = 0, all = T)
IBET.48h.camera <- IBET.48h.camera %>% filter(Direction.x == "Up")

# convert FDR to -log10(FDR)
IBET.24h.camera$LogFDR = -log10(IBET.24h.camera$FDR.x)
IBET.48h.camera$LogFDR = -log10(IBET.48h.camera$FDR.x)

# dotplot
IBET.24h.camera$Timepoint = "IBET.24h"
IBET.48h.camera$Timepoint = "IBET.48h"

all.samps <- rbind(IBET.24h.camera, IBET.48h.camera)
dat <- all.samps
dat$RandomIndex <- sample(1:nrow(dat), nrow(dat))
alphabetical <- IBET.24h.camera[order(IBET.24h.camera$Row.names),]
dat$alphabetical <- seq(1,nrow(dat))
dat <- dat[order(-dat$LogFDR), ]
pos <- dat[dat$Direction.x == "Up", ]
pos <- pos %>% mutate(color = if_else(pos$LogFDR>2,true = "Sig", false = "NotSig"))
neg <- dat[dat$Direction.x == "Down", ]

dat <- pos %>% mutate(color = if_else(pos$LogFDR>2,true = "Sig", false = "NotSig"))
dat$label <- as.character(dat$Row.names)
dat$label <- ifelse(dat$LogFDR>2 & dat$PropUp > 0.3 | dat$Row.names == "SF3B3" | dat$Row.names == "SF3B1", dat$label, "")
ixlabel <- which(dat$LogFDR>2)
dat$label[ixlabel] <- dat$Row.names[ixlabel] 

# Dotplot
ggplot(dat, aes(RandomIndex, LogFDR, label = label, size = PropUp, col = color)) + 
  geom_point(alpha = 0.4) +
  scale_y_continuous(trans = "log1p") +
  scale_colour_manual(values = c("grey60", "green3")) +
  scale_size_continuous(breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1)) +
  theme_bw() +
  #ylim(c(0,3)) +
  geom_hline(yintercept = 2, color = "black", linetype = 3) +
  facet_wrap(~ Timepoint) + 
  ggrepel::geom_text_repel(size = 4, max.overlaps = 1000)
ggsave(filename = "./IBET_EdgeR_dotplots.pdf", width = 5.5, height = 3.5)
```

## 24 vs 48hr scatterplot
```{r}
dat %>% select(Row.names, LogFDR, Timepoint) %>% pivot_wider(values_from = LogFDR, names_from = Row.names, id_cols = Timepoint) %>% column_to_rownames("Timepoint") %>% t() %>% as.data.frame() %>% rownames_to_column() %>% as.tibble() %>% 
  mutate(labs = ifelse(IBET.24h > 2 | IBET.48h > 2, rowname, "")) %>%
  mutate(sig = ifelse(IBET.24h > 2 & IBET.48h > 2, "sig.both", ifelse(IBET.24h > 2 & IBET.48h < 2, "sig.24", 
                             ifelse(IBET.24h < 2 & IBET.48h > 2, "sig.48", "not.sig")))) %>%
  ggplot(aes(x = IBET.24h, y = IBET.48h, label = labs, color = sig)) + geom_point() +
  scale_x_continuous(trans = "log1p") +
  scale_y_continuous(trans = "log1p") +
  geom_hline(yintercept = 2, linetype = 3) +
  geom_vline(xintercept = 2, linetype = 3) +
  ggrepel::geom_text_repel() + theme_classic() +
  scale_color_manual(values = c("grey40", "red", "blue", "green"))
ggsave(filename = "edgeR/plots/CHROM_IBET_edgeR_dotplot_24_v_48.pdf", width = 4, height = 4)
```

## Barcode plots for hits
```{r}
pdf(file.path(plots.dir, "CHROM_IBET_barcode_plots.pdf"), useDingbats = F)
lapply(hits, function(x){
  barcodeplot(qlf.D1$table$logFC, 
              index = genesymbollist[[as.character(x)]], 
              main = paste("Barcodeplot for", as.character(x)),
              labels = c("Negative logFC", "Positive logFC"), quantile = c(-0.5, 0.5))
})
dev.off()
```

## count plots
```{r}
counts <- cbind(dge$genes, cpm(dge$counts))

# SF3B3
pdf("edgeR/plots/SF3B3_IBET_counts.pdf", width = 3, height = 3.5)

counts <- cbind(dge$genes, cpm(dge$counts))
counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R1", "CHROM.24h.IBET.SORT.R1")) %>% filter(Symbol == "SF3B3") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R1", "CHROM.48h.IBET.SORT.R1")) %>% filter(Symbol == "SF3B3") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R2", "CHROM.24h.IBET.SORT.R2")) %>% filter(Symbol == "SF3B3") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R2", "CHROM.48h.IBET.SORT.R2")) %>% filter(Symbol == "SF3B3") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

dev.off()

# ARID1A
pdf("edgeR/plots/ARID1A_IBET_counts.pdf", width = 3, height = 3.5)

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R1", "CHROM.24h.IBET.SORT.R1")) %>% filter(Symbol == "ARID1A") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R1", "CHROM.48h.IBET.SORT.R1")) %>% filter(Symbol == "ARID1A") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R2", "CHROM.24h.IBET.SORT.R2")) %>% filter(Symbol == "ARID1A") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R2", "CHROM.48h.IBET.SORT.R2")) %>% filter(Symbol == "ARID1A") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

dev.off()

# SF3B1
pdf("edgeR/plots/SF3B1_IBET_counts.pdf", width = 3, height = 3.5)

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R1", "CHROM.24h.IBET.SORT.R1")) %>% filter(Symbol == "SF3B1") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R1", "CHROM.48h.IBET.SORT.R1")) %>% filter(Symbol == "SF3B1") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R2", "CHROM.24h.IBET.SORT.R2")) %>% filter(Symbol == "SF3B1") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R2", "CHROM.48h.IBET.SORT.R2")) %>% filter(Symbol == "SF3B1") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

dev.off()

# ASH2L
pdf("edgeR/plots/ASH2L_IBET_counts.pdf", width = 3, height = 3.5)

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R1", "CHROM.24h.IBET.SORT.R1")) %>% filter(Symbol == "ASH2L") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R1", "CHROM.48h.IBET.SORT.R1")) %>% filter(Symbol == "ASH2L") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.24h.IBET.LIB.R2", "CHROM.24h.IBET.SORT.R2")) %>% filter(Symbol == "ASH2L") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

counts %>% dplyr::select(c("Symbol","Guide", "CHROM.48h.IBET.LIB.R2", "CHROM.48h.IBET.SORT.R2")) %>% filter(Symbol == "ASH2L") %>% reshape2::melt() %>% ggplot(aes(variable, value, color = Guide, group = Guide)) + geom_line() + geom_point(size = 3) + scale_y_continuous(trans = "log10") + theme_bw()

dev.off()


```

## Rank plots - combined timepoints analysis
```{r}
# D1 group
dat <- qlf.D1$table %>% 
  rownames_to_column() %>% 
  mutate(gene = str_split(rowname, "_", simplify = T)[,1]) %>%
  group_by(gene) %>% 
  summarise(median(logFC)) %>% 
  arrange(desc(`median(logFC)`)) %>% mutate(rank = seq(1,length(gene)))
dat$sig <- ifelse(dat$`median(logFC)` >= quantile(dat$`median(logFC)`, 0.9), "Enriched", ifelse(dat$`median(logFC)` <= quantile(dat$`median(logFC)`, 0.1),"Depleted", "Unchanged"))
dat$label <- ifelse(dat$`median(logFC)` >= quantile(dat$`median(logFC)`, 0.995), dat$gene, ifelse(dat$`median(logFC)` <= quantile(dat$`median(logFC)`, 0.005),dat$gene, ""))

ggplot(dat) + 
  geom_point(aes(x = rank, y = `median(logFC)`, color = sig, label = label)) + 
  theme_linedraw() + 
  scale_color_manual(values = c("blue", "red", "grey80")) + 
  ggrepel::geom_text_repel(aes(x = rank, y = `median(logFC)`, label = label), median.overlaps = 100) +
  labs(title = "CHROM IBET - D1 - SORT fraction - Rank plot", y = "median Log2FC", x = "Gene")
  #coord_cartesian(ylim = c(-2,4))
ggsave(filename = "edgeR/plots/CHROM_IBET_D1_Rank_plot_median_logFC.pdf", width = 3, height = 4)

# D2 group
dat <- qlf.D2$table %>% 
  rownames_to_column() %>% 
  mutate(gene = str_split(rowname, "_", simplify = T)[,1]) %>%
  group_by(gene) %>% 
  summarise(median(logFC)) %>% 
  arrange(desc(`median(logFC)`)) %>% mutate(rank = seq(1,length(gene)))
dat$sig <- ifelse(dat$`median(logFC)` >= quantile(dat$`median(logFC)`, 0.9), "Enriched", ifelse(dat$`median(logFC)` <= quantile(dat$`median(logFC)`, 0.1),"Depleted", "Unchanged"))
dat$label <- ifelse(dat$`median(logFC)` >= quantile(dat$`median(logFC)`, 0.995), dat$gene, ifelse(dat$`median(logFC)` <= quantile(dat$`median(logFC)`, 0.005),dat$gene, ""))

ggplot(dat) + 
  geom_point(aes(x = rank, y = `median(logFC)`, color = sig, label = label)) + 
  theme_linedraw() + 
  scale_color_manual(values = c("blue", "red", "grey80")) + 
  ggrepel::geom_text_repel(aes(x = rank, y = `median(logFC)`, label = label), median.overlaps = 100) +
  labs(title = "CHROM IBET - D2 - SORT fraction - Rank plot", y = "median Log2FC", x = "Gene")
  #coord_cartesian(ylim = c(-2,4))
ggsave(filename = "edgeR/plots/CHROM_IBET_D2_Rank_plot_median_logFC.pdf", width = 3, height = 4)

```

## Session info
```{r}
sessionInfo()
```
